---
title: "TidyTuesday: Urban Green Space"
page-layout: full
toc: false
---

```{r}
# Load required libraries
library(tidyverse)
```

```{r}
urban <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-09-22/urban.csv")
```


::: {.tt-hero}
# 🌿 How Green Are the World's Cities?

### Exploring urban green space through three connected visualizations

For this TidyTuesday project, I explored how green space is distributed
across cities, how it differs across world regions, and whether cities with
more green area also provide more green space per person.

:::


```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)

urban <- read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-09-22/urban.csv"
)

urban_2025 <- urban |>
  filter(year == 2025)
```


## About the Data

:::: {.tt-facts}

::: {.tt-fact .tt-pink}

### 🌎 Source

UN-Habitat  
Urban Indicators Dataset

:::

::: {.tt-fact .tt-blue}

### 📍 Observations

Cities and urban areas  
across different years

:::

::: {.tt-fact .tt-green}

### 📊 Size

**5,600 observations**  
**10 variables**

:::

::::


The dataset includes measurements such as the percentage of urban land
covered by green area and the amount of green space available per person.


---

# 01. Viz

::: {.tt-question}

### Question

**How is the percentage of green area distributed across cities around the world in 2025?**

:::


```{r}
urban_2025 <- urban |>
  filter(year == 2025)

ggplot(urban_2025, aes(x= averageShareOfGreenAreaInCityUrbanAreaPct))+
  geom_histogram(fill = "darkgreen", color = "white")+
  labs(title = "Distribution of Green Area Across Cities in 2025", 
       x = "Average share of green area (%)", 
       y = "Number of cities"
       )+
  theme_minimal()
```



:::: {.columns}

::: {.column width="45%"}

::: {.tt-insight .tt-green-border}

### 💡 What I Found

Among cities with available data, many are concentrated around roughly
**3%–10% green area**.

The distribution is right-skewed, meaning cities with especially high
percentages of green space are much less common.

:::

:::

::: {.column width="40%"}

::: {.tt-mini}

### Why this works

- Makes the overall distribution easy to see
- Clearly shows where most cities are concentrated

### What I would improve

- Experiment with bin width
- Add a reference line or annotation

:::

:::

::::


---

# 02. Viz

::: {.tt-question}

### Question

**Do cities in different world regions have different amounts of green area?**

:::


```{r}
ggplot(urban_2025, aes(x = averageShareOfGreenAreaInCityUrbanAreaPct, y = sdgRegion)) +
  geom_boxplot(fill = "darkblue") +
  labs(title = "Green Area Across World Regions", x = "% of green area", y = "region") +
  theme()
```


:::: {.columns}

::: {.column width="50%"}

::: {.tt-insight .tt-blue-border}

### 💡 What I Found

Green-area percentages vary across world regions.

Some regions show relatively higher values, while others tend to have
lower percentages. There is also substantial variation **within** many
regions, meaning cities in the same broad region can still look very different.

:::

:::

::: {.column width="40%"}

::: {.tt-mini}

### Why this works

- Horizontal layout makes region labels easier to read
- Boxplots show both typical values and variation

### What I would improve

- Reorder regions for easier comparison
- Clean up long region labels

:::

:::

::::


---

# 03. Viz

::: {.tt-question}

### Question

**Do cities with more green area also have more green space per person?**

:::


```{r}
ggplot(urban_2025, aes(x = averageShareOfGreenAreaInCityUrbanAreaPct, y = greenAreaPerCapitaM2)) +
  geom_point(color = "darkblue", alpha = 0.5) +
  labs(title = "Does More Green Area Mean More Green Space per Person?",
       x = "% of Green Area",
       y = "Green area per capita"
       )
```


:::: {.columns}

::: {.column width="40%"}

::: {.tt-insight .tt-purple-border}

### 💡 What I Found

Cities with higher percentages of green area generally also tend to provide
more green space per person.

However, the relationship is not perfect. The scatterplot still contains
substantial variation and several outliers.

:::

:::

::: {.column width="50%"}

::: {.tt-mini}

### Why this works

- Directly compares two quantitative variables
- Transparency helps reduce overlapping points

### What I would improve

- Label interesting outliers
- Explore whether population size explains some variation

:::

:::

::::


---

::: {.tt-takeaway}

# 🌱 Final Takeaway

These three visualizations tell a connected story.

**First**, high percentages of urban green space are relatively uncommon.

**Second**, access to green space differs across world regions.

**Finally**, cities with more green area generally also offer more green
space per person — but there are important exceptions.

This project helped me move from a **univariate distribution** to
**regional comparison** and finally to a **relationship between two
quantitative variables**, while using visualization to build a connected
data story.

:::
